---
title: Penalized Regression Practice
output: pdf_document
---

# Installing packages
```{r}
library(MASS)
library(tidyverse)
library(caret)
library(glmnet)
library(caTools)
```

# Reading data
```{r}
data("Boston", package = "MASS")
str(Boston)
```

# Split training and test data
```{r}
set.seed(123)
training_samples <- Boston$medv %>%
  createDataPartition(p=0.75,list=FALSE)
train.data <- Boston[training_samples,]
test.data <- Boston[-training_samples,]
nrow(train.data)
nrow(test.data)
```

# Ridge regression model - finding best lambda
```{r}
x <- model.matrix(medv~.,train.data)[,-1]
y <- train.data$medv
cv <- cv.glmnet(x,y,alpha=0)
cv$lambda.min
```
Best lambda for ridge regression is 0.6490823

# Ridge regression model - coefficient of the fitted model
```{r}
model <- glmnet(x,y,alpha=0,lambda=cv$lambda.min)
coef(model)
```
# Ridge regression model - prediction, RMSE and R^2
```{r}
x.test <- model.matrix(medv~., test.data)[,-1]
predictions <- model %>% predict(x.test) %>% as.vector()
data.frame(
  RMSE = RMSE(predictions,test.data$medv),
  Rsquare = R2(predictions, test.data$medv)
)
```

# Lasso regression model - finding best lambda
```{r}
cv <- cv.glmnet(x,y,alpha=1)
cv$lambda.min
```

# Lasso regression model - coefficients of fitted model
```{r}
model <- glmnet(x,y,alph=1,lambda=cv$lambda.min)
coef(model)
```

# Lasso regression model - prediction, RMSE and R^2
```{r}
x.test <- model.matrix(medv~., test.data)[,-1]
predictions <- model %>% predict(x.test) %>% as.vector()
data.frame(
  RMSE=RMSE(predictions,test.data$medv),
  Rsquare=R2(predictions,test.data$medv)
)
```

# Elastic net model - best tuning parameter 
```{r}
model <- train(medv~., data=train.data, method="glmnet", trControl = trainControl("cv",number=10),tuneLength=10)
model$bestTune
```
# Elastic net model - coefficients of fitted model
```{r}
coef(model$finalModel, model$bestTune$lambda)
```

# Elastic net model - prediction, RMSE and R^2
```{r}
x.test <- model.matrix(medv~., test.data)[,-1]
predictions <- model %>% predict(x.test)
data.frame(
  RMSE = RMSE(predictions, test.data$medv),
  Rsquare = R2(predictions, test.data$medv)
)
```